---
title: "Transient Detection"
author: "Hooker Lab"
output: html_document
---

```{r rmd setup,include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning=FALSE)
```

```{r r setup,message=F,cache=F}
library(here)
source(here::here("simulation_fns.R"))
```

# Algorithm

The simple transient detection algorithm starts by calculating the Euclidean distance from the final observation to all other ones. The earliest time index within the closest $x \%$ is chosen as the cutoff between transient and limit cycle. We pick $2\%$ as default. This procedure can use the spline-smoothed position or (currently noiseless) gradient.

```{r get_cutoff_fn,echo=TRUE}
get_lc_cutoff <- function(full_samples,cutoff_prop=50,gradient=FALSE){
  
  # this naive algorithm delineates the limit cycle from transient
  # it first ranks the nearest neighbors in the entire signal 
  # the earliest nearest neighbor in the closest `1/cutoff_prop`-fraction
  # is picked to start the signal 
  
  x_var <- ifelse(gradient,"f_x","x")
  y_var <- ifelse(gradient,"f_y","y")
  
  final_observation <- full_samples[nrow(full_samples),c(x_var,y_var)]
  other_observations <- full_samples[-nrow(full_samples),c(x_var,y_var)]
  
  cutoff_index <- min(RANN::nn2(other_observations,final_observation,
    k=nrow(other_observations))$nn.idx[1:(nrow(other_observations)/cutoff_prop)])
  
  return(cutoff_index)
}
```

```{r plotting_fns}

get_trajectory_plot <- function(full_samples,transient_cutoff,gradient=FALSE){

  x_var <- ifelse(gradient,"f_x","x")
  y_var <- ifelse(gradient,"f_y","y")

  # add indicator if in transient cutoff range
  full_samples_toplot <- full_samples %>% mutate(transient = ifelse(row_number() <= transient_cutoff,"Transient","Limit Cycle"))

  full_traj <- ggplot2::ggplot(full_samples_toplot,aes(x=.data[[x_var]],y=.data[[y_var]])) +
    geom_point(color = "#FDB515", alpha = 0.7) +
    facet_wrap(~transient) +
    labs(x = x_var, y = y_var, title = paste0("Trajectory with Cutoff at ",transient_cutoff, " Samples"))

  return(full_traj)
}

get_axis_plots <- function(full_samples,transient_cutoff,gradient=FALSE){

  x_var <- ifelse(gradient,"f_x","x")
  y_var <- ifelse(gradient,"f_y","y")

  full_samples_toplot <- full_samples %>%
    mutate(index = row_number()) %>%
    select(!!x_var,!!y_var,index) %>%
    pivot_longer(-index,names_to="axis",values_to="position")

  axis_plots <- ggplot2::ggplot(full_samples_toplot,aes(x=index,y=position)) +
    geom_point(alpha=0.7) +
    geom_vline(xintercept = transient_cutoff, linetype = "longdash")+
    facet_wrap(~axis) +
    labs(x = "Sample Index", y = "Position", title = "Projection onto Each Axis")

  return(axis_plots)
}
```


# Van der Pol Examples

The classic Van der Pol with $\mu = 1.5$ is used to demonstrate robustness to the sampling rate and fraction of closest neighbors considered.

## Default: 100 Hz and Top 0.02

```{r demo_default}

sample_deltaT <- 0.01 # inverse of sampling rate
vdp_data <- list(name = "VDP", system = "van_der_pol", params = list(mu=1.5),
                              n = 5000, lc_tail_n = 5000, sample_density = sample_deltaT, var_x = 0.05, var_y = 0.05,
                              x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, 
                              smoother = "bspline", data_seed = 2, noise_seed = 2)
experiment_list <- list(vdp_data)
experiment_data <- generate_data_object_model(experiment_list)

cutoff_index <- get_lc_cutoff(experiment_data[[1]]$limit_cycle_samples)

full_samples <- experiment_data[[1]]$limit_cycle_samples
trajectories <- get_trajectory_plot(full_samples,cutoff_index,gradient=F)
projections <- get_axis_plots(full_samples,cutoff_index,gradient=F)
print(trajectories)
print(projections)
```

## 20 Hz and Top 0.02

### Position

```{r}

sample_deltaT <- 0.05 # inverse of sampling rate
vdp_data <- list(name = "VDP", system = "van_der_pol", params = list(mu=1.5),
                              n = 5000, lc_tail_n = 5000, sample_density = sample_deltaT, var_x = 0.05, var_y = 0.05,
                              x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, 
                              smoother = "bspline", data_seed = 2, noise_seed = 2)
experiment_list <- list(vdp_data)
experiment_data <- generate_data_object_model(experiment_list)

cutoff_index <- get_lc_cutoff(experiment_data[[1]]$limit_cycle_samples)

full_samples <- experiment_data[[1]]$limit_cycle_samples
trajectories <- get_trajectory_plot(full_samples,cutoff_index,gradient=F)
projections <- get_axis_plots(full_samples,cutoff_index,gradient=F)
print(trajectories)
print(projections)
```

### Gradient

```{r}

sample_deltaT <- 0.05 # inverse of sampling rate
vdp_data <- list(name = "VDP", system = "van_der_pol", params = list(mu=1.5),
                              n = 5000, lc_tail_n = 5000, sample_density = sample_deltaT, var_x = 0.05, var_y = 0.05,
                              x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, 
                              smoother = "bspline", data_seed = 2, noise_seed = 2)
experiment_list <- list(vdp_data)
experiment_data <- generate_data_object_model(experiment_list)

cutoff_index <- get_lc_cutoff(experiment_data[[1]]$limit_cycle_samples)

full_samples <- experiment_data[[1]]$limit_cycle_samples
trajectories <- get_trajectory_plot(full_samples,cutoff_index,gradient=T)
projections <- get_axis_plots(full_samples,cutoff_index,gradient=T)
print(trajectories)
print(projections)
```

## 50 Hz and Top 0.02

### 1000 Samples

```{r}

sample_deltaT <- 0.1 # inverse of sampling rate
vdp_data <- list(name = "VDP", system = "van_der_pol", params = list(mu=1.5),
                              n = 1000, lc_tail_n = 1000, sample_density = sample_deltaT, var_x = 0.05, var_y = 0.05,
                              x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, 
                              smoother = "bspline", data_seed = 2, noise_seed = 2)
experiment_list <- list(vdp_data)
experiment_data <- generate_data_object_model(experiment_list)

cutoff_index <- get_lc_cutoff(experiment_data[[1]]$limit_cycle_samples)

full_samples <- experiment_data[[1]]$limit_cycle_samples
trajectories <- get_trajectory_plot(full_samples,cutoff_index,gradient=F)
projections <- get_axis_plots(full_samples,cutoff_index,gradient=F)
print(trajectories)
print(projections)
```

### 5000 Samples

```{r}

sample_deltaT <- 0.1 # inverse of sampling rate
vdp_data <- list(name = "VDP", system = "van_der_pol", params = list(mu=1.5),
                              n = 5000, lc_tail_n = 5000, sample_density = sample_deltaT, var_x = 0.05, var_y = 0.05,
                              x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, 
                              smoother = "bspline", data_seed = 2, noise_seed = 2)
experiment_list <- list(vdp_data)
experiment_data <- generate_data_object_model(experiment_list)

cutoff_index <- get_lc_cutoff(experiment_data[[1]]$limit_cycle_samples)

full_samples <- experiment_data[[1]]$limit_cycle_samples
trajectories <- get_trajectory_plot(full_samples,cutoff_index,gradient=F)
projections <- get_axis_plots(full_samples,cutoff_index,gradient=F)
print(trajectories)
print(projections)
```

## 100 Hz and Top 0.5

```{r}

sample_deltaT <- 0.01 # inverse of sampling rate
vdp_data <- list(name = "VDP", system = "van_der_pol", params = list(mu=1.5),
                              n = 5000, lc_tail_n = 5000, sample_density = sample_deltaT, var_x = 0.05, var_y = 0.05,
                              x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, 
                              smoother = "bspline", data_seed = 2, noise_seed = 2)
experiment_list <- list(vdp_data)
experiment_data <- generate_data_object_model(experiment_list)

cutoff_index <- get_lc_cutoff(experiment_data[[1]]$limit_cycle_samples,
                              cutoff_prop=2)

full_samples <- experiment_data[[1]]$limit_cycle_samples
trajectories <- get_trajectory_plot(full_samples,cutoff_index,gradient=F)
projections <- get_axis_plots(full_samples,cutoff_index,gradient=F)
print(trajectories)
print(projections)
```

## 100 Hz and Top 0.95

```{r}

sample_deltaT <- 0.01 # inverse of sampling rate
vdp_data <- list(name = "VDP", system = "van_der_pol", params = list(mu=1.5),
                              n = 5000, lc_tail_n = 5000, sample_density = sample_deltaT, var_x = 0.05, var_y = 0.05,
                              x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, 
                              smoother = "bspline", data_seed = 2, noise_seed = 2)
experiment_list <- list(vdp_data)
experiment_data <- generate_data_object_model(experiment_list)

cutoff_index <- get_lc_cutoff(experiment_data[[1]]$limit_cycle_samples,
                              cutoff_prop=1.05)

full_samples <- experiment_data[[1]]$limit_cycle_samples
trajectories <- get_trajectory_plot(full_samples,cutoff_index,gradient=F)
projections <- get_axis_plots(full_samples,cutoff_index,gradient=F)
print(trajectories)
print(projections)
```

# Rosenzweig-MacArthur

## 100 Hz and Top 0.02 with Smoothing

### Position

```{r}
rz_model <- list(name = "RZ", system = "rzma", params = list(a = 2.8, b = 0.7, c = 1.35, r = 3.5, K = 1.4), n = 1000, sample_density = 0.1, var_x = 0.03, var_y = 0.03,
                    lc_tail_n = 1000, x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, smoother = "bspline", data_seed = 1, noise_seed = 2)
experiment_list <- list(rz_model)
experiment_data <- generate_data_object_model(experiment_list)

cutoff_index <- get_lc_cutoff(experiment_data[[1]]$limit_cycle_samples,
                              cutoff_prop=50)

full_samples <- experiment_data[[1]]$limit_cycle_samples
trajectories <- get_trajectory_plot(full_samples,cutoff_index,gradient=F)
projections <- get_axis_plots(full_samples,cutoff_index,gradient=F)
print(trajectories)
print(projections)
```


### Gradient

```{r}
rz_model <- list(name = "RZ", system = "rzma", params = list(a = 2.8, b = 0.7, c = 1.35, r = 3.5, K = 1.4), n = 1000, sample_density = 0.1, var_x = 0.03, var_y = 0.03,
                    lc_tail_n = 1000, x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, smoother = "bspline", data_seed = 1, noise_seed = 2)
experiment_list <- list(rz_model)
experiment_data <- generate_data_object_model(experiment_list)

cutoff_index <- get_lc_cutoff(experiment_data[[1]]$limit_cycle_samples,
                              cutoff_prop=50)

full_samples <- experiment_data[[1]]$limit_cycle_samples
trajectories <- get_trajectory_plot(full_samples,cutoff_index,gradient=T)
projections <- get_axis_plots(full_samples,cutoff_index,gradient=T)
print(trajectories)
print(projections)
```


## 100 Hz and Top 0.02 without Smoothing

```{r}
rz_model <- list(name = "RZ", system = "rzma", params = list(a = 2.8, b = 0.7, c = 1.35, r = 3.5, K = 1.4), n = 1000, sample_density = 0.1, var_x = 0.03, var_y = 0.03,
                    lc_tail_n = 1000, x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, smoother = "orig", data_seed = 1, noise_seed = 2)
experiment_list <- list(rz_model)
experiment_data <- generate_data_object_model(experiment_list)

cutoff_index <- get_lc_cutoff(experiment_data[[1]]$limit_cycle_samples,
                              cutoff_prop=50)

full_samples <- experiment_data[[1]]$limit_cycle_samples[,1:4]
trajectories <- get_trajectory_plot(full_samples,cutoff_index,gradient=F)
projections <- get_axis_plots(full_samples,cutoff_index,gradient=F)
print(trajectories)
print(projections)
```