---
title: "Van der Pol Test Environment"
author: "William Torous"
date: "`r Sys.Date()`"
output: html_document
---

This markdown file is a minimum reproducible example of the simulation pipeline using a Van der Pol system.

```{r rmd setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE)
```

```{r r setup,message=F,cache=F}
library(here)
source(here::here("simulation_fns.R"))
```

## Data Generation

500 samples from the tail of the limit cycle are generated.

```{r get data}
some_noise <- list(name = "VDP", system = "van_der_pol", params = list(mu=3), n = 1000, sample_density = 0.1, var_x = 0.05, var_y = 0.05,
                   lc_tail_n = 500, x_grid_size = 36, y_grid_size = 36, extrapolation_size = 0.5, smoother = "bspline", data_seed = 2, noise_seed = 2)
experiment_list <- list(some_noise)
experiment_data <- generate_data_object(experiment_list)
```

## Estimator Specification

The estimators which the simulations compare are defined here.

```{r define esimtators}
truth <- list(method = "truth",  params = list())
spline <- list(method = "spline",  params = list(lambda = 1e-2, norder = 6, nbasis = 20, side_info = list()))
```

## Fit Estimators

Here the estimators are fit, but not analyzed.

```{r fit estimators}
experiment_estimators <- list(truth, spline)
experiment_results <- evaluate_gradient_methods(experiment_data, experiment_estimators)
```

## Analyze Results

Plots are used to compare the various estimators.

```{r analysis}
plot_delta <- list(type = "field_delta_paths", experiments = list(c(data = 1, estimator = 1, ref = 1),c(data = 1, estimator = 2, ref = 1)))
visualize_results(experiment_results, list(plot_delta))
```



